---
title: "Practical ML Course"
output: html_document
---


###Goal
Predict the "classe" variable for a set of test observations

```{r}
library(caret)
setwd("/Users/svetvaz/projects/coursera/Practical_ML/")
trainingfile<- read.table("./pml-training.csv",sep=",",na.strings = c("NA",""),header=TRUE)
testingfile<- read.table("./pml-testing.csv",sep=",",na.strings = c("NA",""),header=TRUE)

intrain<-createDataPartition(trainingfile$classe,p=0.75,list=FALSE)
training<-trainingfile[intrain,]
training_crossval<-trainingfile[-intrain,]

#remove all the columns with missing values
training<-training[,colSums(is.na(training))==0]
classe<-training$classe
isnumeric<-sapply(training,is.numeric)
training<-cbind(classe,training[,isnumeric])
training$X<-training$num_window<-NULL

training_crossval<-training_crossval[,colSums(is.na(training_crossval)) == 0]
vclasse<-training_crossval$classe
training_crossval_is_numeric <- sapply(training_crossval, is.numeric)
training_crossval<-cbind(vclasse,training_crossval[,training_crossval_is_numeric])
colnames(training_crossval)[1]<-"classe"
training_crossval$X<-training_crossval$num_window<-NULL


testingfile<-testingfile[,colSums(is.na(testingfile)) == 0]
tnums <- sapply(testingfile, is.numeric)
testingfile<-testingfile[,tnums]
testingfile$X<-testingfile$num_window<-NULL
```

###Creating the model
The model is being built using the random forest algorithm (with default parameters) which is the best in accuracy amongst current algorithms for classification.It has an effective method for estimating missing data and maintains accuracy when a large proportion of the data are missing. The "classe" variable is the target variable and all other variables from the training set are used as features
```{r}
model <- train(training$classe~.,data=training, method="rf")
model$results
```

###Out of Sample Error Rate Calculation
Using a cross validation fold of 8, the cross validation set that was previously obtained from the training set is evaluated
```{r}
trainingCV <- trainControl(method = "cv", number = 8)
model_CV <- train(training_crossval$classe~.,data=training_crossval, method="rf",trControl=trainingCV)
model_CV$results
confusionMatrix(predict(model_CV, newdata=training_crossval), training_crossval$classe)
model_CV$finalModel
```

### Running the model on the testing data
Finally the model created above is evaluated on the 20 test cases and the predictions are written to a file
```{r}
predictions <- predict(model, testingfile)
print (predictions)


pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predictions)

```